Fault-tolerant Asynchronous Circuits

### 2019-12-04 draft_final.tex
edited first half of ch 2, not lit review

\chapter{Related Works}
\section{Faults}
Faults can be split into permanent and non-permanent faults. Non-permanent faults are further split into transient and intermittent faults \cite{jha_gupta_2003}. %copied 
A fault is defined as the physical difference between the good or correct system and the current system. An error may be caused by a fault, which is when the state of the system differs from the state of the good system and thus unable to perform some specified function. Hidden faults are also possible where the fault does not cause an error \cite{jha_gupta_2003}. \\

Permanent faults refer to the presence of a fault that affects the functional behavior of a system(circuit) permanently. Some examples that may cause permanent faults are: incorrect connections between integrated circuits, broken components or parts of components, incorrect silicon and metal connections (manufacturing problem), and functional design errors. Most permanent faults occur during fabrication. There are many existing literatures that cover permanent faults and the different tests at production time that can uncover the presence and location of these faults. \cite{jha_gupta_2003} \cite{giz_book_2006}. We ignore the possible manufacturing problems and assume that the circuit is free from those problems. Instead we consider the permanent faults that develop during operation from radiation. For example, high energy ions can impact the transistors and the circuit causing short-circuiting, rupture of the gate oxide insulation and other permanent faults.\\ %copied from SEE eetimes article

Non-permanent faults are defined as faults that affect the system's functional behavior for a finite but unknown period of time. They can occur at random moments and are categorized into two types, intermittent and transient faults. Intermittent faults are caused by non-environmental conditions such as loose connections, deteriorating or aging components, critical timing (hazards and race conditions), resistance and capacitance variations (which may lead to timing faults), physical irregularities, and noise. Intermittent faults usually affect the system for a shorter time period than the application time of a test developed for permanent faults and thus may not be detected using such a test. Intermittent faults can transition into permanent faults naturally but may take a long time from hours up to months. \\

Transient faults are caused by environmental conditions such as cosmic rays, $\alpha$-particles, pollution, humidity, temperature, pressure, vibration, power supply fluctuations, electromagnetic interference, static electrical discharges, and ground loops. In particular $\alpha$-particles are a major cause of these faults. These $\alpha$-particles can interact with the semiconductor material which results in the generation of electron-hole pairs. When these $\alpha$-particles collide and interact with the circuit it can induce a voltage/current pulse. Also known as single-event transients (SETs), if the width of this pulse or transient is wide enough it may propagate through the circuit. $\alpha$-particles may also interact with a memory element (of which flip flops, latches and C-elements are most relevant to this work) and cause a bit flip in the state of the memory element. This is also known as single-event upsets (SEUs), SEUs include both single-bit and multi-bit upsets but single-bit upsets are most common and much work has been done to eradicate failures from SEUs. 
Transient faults may be hard to detect because it appears sporadically and may not affect the system during testing. \\

In this work we will not be concerned with intermittent faults since it can be thought of as a transition between transient to permanent faults. Instead the two main fault modes we will look into are transient and permanent faults. In both radiation effects, also called single event effects (SEEs), figure prominently. SEEs are an umbrella term that covers the aforementioned SETs and SEUs as well as the single event effects causing permanent faults. There are two main sources of radiation, one is cosmic rays from space which interact with the atmosphere to produce high energy neutrons. Another is that $\alpha$-particles can be caused by trace impurities of radioactive elements present in the packaging material of the chip and the silicon substrate. Radiation from cosmic rays become more severe at higher elevations while the latent source in packaging is difficult to eliminate. Thus, initially a lot of the work in fault-tolerant design was done for aviation and space flight.

@@

\section{Asynchronous Circuit and Synthesis}
In the previous chapter we described how an asynchronous circuit works from a higher-level perspective and in particular that it should work under varying delays. In this section we will cover some definitions and properties of the circuit and also how we can synthesize the circuit from a higher-level specification. Both Speed Independent and Delay Insensitive technologies will be introduced. There are differences but also many similarities between the two methods. Even though this work is based on the Speed Independent model, there has been some previous work done on the Quasi-delay Insensitive model so I wanted to provide some context to make a direct comparison.

\subsection{Speed Independent Circuits}
As mentioned in the introduction, circuits are composed of gates and wires. The unbounded gate delay model assumes an arbitrary delay on all the gates and no delay on wires. This delay model is the basis for Muller circuits \cite{Muller_59}. Circuits that operate correctly regardless of the delay are called Speed Independent Circuits. The correct operation is specified when designing the circuit, for example, given some input values the circuit will return specified output values. 
There are many ways to represent specifications such as Petri nets and state graphs, and there are established methods to build asynchronous circuits to satisfy the specifications. We won't go into these representations and methods in detail, other than the aspects that are relevant to this work. We consider circuits which has wires that take on Boolean values of either `1' or `0'.\\
%Wires in the circuits take on Boolean values of either `1' or `0'. A state is defined by assigning a Boolean value to each wire of the circuit. States are unique in that no two states share the same wire values. There will be a formal definition of wires and states in the next chapter.

@@

%C-elements
The C-element proposed by Muller is an important building block for asynchronous circuits. It acts as a memory element in that it receives two inputs. When both inputs are low, the C-element output goes low, and when both inputs are high the C-element output goes high. However, when the inputs are mismatched then the output gets held at its previous value. This can be achieved in a number of ways such as a gate implementation as well as a transistor level implementation with feedback using a weak transistor, other implementations use capacitance to achieve this weak feedback.
\begin{figure}
\includegraphics[width=\textwidth]{celement}
\captionof{figure}[C-element: logic and implementation]{a. logic of the C-element. A, B are inputs, C is the output \\ b. diagram of a C-element at the gate level c. transistor-level implementation}
\end{figure}

C-elements can be used in various asynchronous circuit designs and is an essential component of Muller circuits. \\

The two specific Muller circuit implementations that we consider are the generalized C-element and standard C-implementations. We assume that the circuits we want to make fault-tolerant are synthesized according to these two implementations. 
Each wire in the circuit is the output wire of a gate logic. The gate logic has a set of input wires and the output wire changes value according to the values of the inputs. We look at two conditions, set, the condition for the output wire to transition to a high value; and reset, the condition for the output wire to transition to a low value. The gate logic then implements the set and reset conditions. 
We skim the details in finding the set and reset conditions from the circuit specifications. For each condition set and reset, there is an ON-set, OFF-set and Don't Care set. ON-set are all input values where the wire is excited and a transition will occur. The OFF-set includes all input values where the wire cannot be excited to some value per the circuit specifications. Input values that are neither in the ON-set or OFF-set are in the Don't Care set. So, for the set condition, the ON-set are all input values where the wire must transition to high. The OFF-Set includes all input values where the wire is excited instead to low, as well as input values for which the output wire must remain low. The input values of the Don't Care set assign the output wire to be either high or low. It does not matter which value is assigned since it does not disturb the function of the circuit and only helps for circuit optimization. \\

The same procedure applies as well in finding the ON-set, OFF-set and Don't Care set for reset except substituting in reset for set and flipping the value of the wire from high to low and vice versa. The set and reset conditions constructed this way then follows the desired circuit specification. The set and reset conditions are implemented using gates and are combined into the output wire through a C-element. This entire block of gates and C-element for the output wire is called the gate logic. \\ %Beerel and Myers.... also expand?

Even though the generalized C-element is presented as a C-element with set and reset as inputs, the transistor level implementation differs slightly. We assume a transistor level implementation of the set and reset combined with the C-element so that there is only one delay for the entire logic block. In addition, this is not a normal C-element because set is used exclusively for the pull-up logic and reset is used for the pull-down logic. In a normal C-element we allow for any four combinations of the values of the set and reset input wires. However, using the pull-up and pull-down logic setup, we do not allow for when both set and reset is `ON' as the output wire would simultaneously be connected to the high voltage and ground. Such an output would have indeterminable voltage values and violate the digital assumptions. Thus, the set and reset logic is designed so that they cannot overlap. This corresponds to $\overline{set}\vee\overline{reset} $ is always TRUE.\\
\begin{center}
\includegraphics[width=.7\textwidth]{genC}
\end{center}
\captionof{figure}[Generalized C-element implementation]{Transistor level implementation of a generalized C-element. The + signs are the set logic inputs and - signs are the reset logic inputs. For this particular element, set is in3 AND in4 while reset is in1 AND in2.}

The main difference is that the set logic and the reset logic must be designed so that they do not turn `ON' at the same time.\\

For the standard C-implementation, we allow for more variation in delay. For each wire in the circuit with a corresponding gate logic, we look at the set and reset logic as composition of various gates (such as AND, OR, NOT). Each of these gates can have their own delay that is arbitrary in length that is instantized as an intermediate wire. Thus, the logic must be synthesized more carefully to eliminate hazards that can occur in these intermediate wires. In particular the set and reset must be activated at states only when the wire is appropriately excited. As a result, the standard C-implementation has a number of set logic functions combined by an OR gate and fed into one input of the C-element. There will also be a number of reset logic functions combined by an OR gate and passed through an inverter and fed into another input of the C-element. This implementation uses the standard C-element so that if the set and reset inputs are mismatched, the output is held. In general, the synthesis can be changed so that there is a hazard-free implementation using only two-input gates.
%explain sop?
\subsection{Quasi Delay Insensitive Circuits}
%QDI
Quasi Delay Insensitive Circuits (QDI) are related to Delay Insensitive Circuits. Previously we mentioned that Delay Insensitive circuits operate under the assumption that both the gate and the wire have unbounded gate delays. In particular each branch of a fork will have different delays and can be treated as separate wires. One result about Delay Insensitive circuits is that it is a very restricted class. Instead if that condition is relaxed so that some of the forks are allowed to be isochronic, then a more general class of circuits can be constructed. Note that in the case where all forks are isochronic, then this becomes the same as speed independent circuits. \\
The gates of a QDI circuit is formed from Production Rules (PRs). For the single output case, a gate takes a set of variables as inputs and outputs a variable. Again, a transition on the variable $x$ is denoted as $x\uparrow$, a transition from $x=0$ to $x=1$; and $x\downarrow$, a transition from $x=1$ to $x=0$. Then the two PRs associated with the gate with output $x$ is :
\begin{align*}
B_u\mapsto x\uparrow \\
B_d\mapsto x\downarrow 
\end{align*}
Where $B_u$ and $B_d$ are called guards. Each guard is a boolean condition formed by some combination of the input variables. When a guard is evaluated to be $True$ then the transition is activated. When it is $False$ then the corresponding transition does not happen. Note that similar to the generalized C-elements, both guards cannot be active at the same time. In addition, it is assumed that each guard is in a sum of products of the input variables. The PRs of a circuit are evaluated all at the same time. At any point in time, if a transition is activated, it will transition in a finite amount of time as long as the corresponding guard is held at $True$. This is due to the assumption of the arbitrary delay of a gate. Note that again, like the generalized C-element implementation, we assume that the transistor level implementation of a PR allows any arbitrary sum of products to be made where there are no delays within the logic of the sum of products.\\

@@

%soft error protection <- error correcting codes for SRAM memory
Because faults cannot be eliminated in digital circuits, different methodologies have been used to make the circuits fault-tolerant. Much of the research done in this area are for synchronous circuits. There has also been some research for asynchronous circuits using some similar techniques. First, we present an overview of some common techniques for synchronous circuits, then introduce the current techniques for asynchronous circuits. Some of the techniques in asynchronous design borrows or are inspired by the much more developed fault-tolerant synchronous circuits field. However due to the difference in timing assumptions of the two types of circuits the techniques are not completely transferable. We describe why some of the solutions from synchronous circuits cannot be directly applied to asynchronous circuits.\\

%- used in the early days for space and flight, and systems that CANNOT go offline/errors like stocks and banking.
%- different techniques... explain a bit about each
%-hardware, software, hybrid, 
%-detection vs correction
%- ECC/ parity checking (needs software/ controller) (Mitra's method)
%-TMR is still super important (possibly why TMR is not viable for async)
In the early days there were certain systems for which eliminating faults and errors were very important. These include space and aviation control systems where due to the altitude there are more radiation and thus more faults. In addition, faults that cause error is very detrimental to these systems and may cause failures leading to loss of life. Other examples of systems requiring reliable circuits include banking and stock market applications where the system must function without error for hours at a time and in case of any error may have devastating consequences, such as a bit flip of the most significant bit in a money transaction. \\

@@

\chapter{Asynchronous Circuit Concepts \& Definitions}
\section{Assumptions:  Boolean signals}
We make the assumption that all signals are Boolean, that is, the signals are either `logic 0' or logic 1'.  When the the gates and C-elements are excited, and their outputs transition, we assume that transition is instantaneous and that after the transition, the signals are still Boolean.   
Even in cases where the wires are affected by faults, we assume that the new value of the wires are Boolean, and if they are inputs into circuit elements, that the output of the circuit elements also remain Boolean.  For each circuit, there is a finite set of wires, which we call $W$.  An exact definition $W$ will be presented at a later section, but for now just think of $W$ as including all the wires of some circuit.

@@

For each element, there exists a function that maps the inputs into a binary value.  This function is represented by $f:[I \to \{0,1\}] \to \{0,1\}$ where $I$ is the set of input wires of the element, a subset of $W$.  If we assume a two input element, we name the input wires $w_1$ and $w_2$.  The function models the designed operation of the element such that if the element had no delay, $f(w_1,w_2)$ is the output value of the element.  With asynchronous delay, if the inputs remain constant, then the output will eventually take the value of $f(w_1,w_2)$.  For gates, such a function is easily defined.  For example, an AND gate such as in figure \ref{fig:gates0} with two inputs $w_1$ and $w_2$ is modeled by the function 
\begin{figure}[h]
\centering
\includegraphics[width=0.4\textwidth]{gates0}
\captionof{figure}{AND gate}
\end{figure}
\[
f_{AND}(w_1,w_2)= w_1\wedge w_2
\]

For C-elements, these are memory holding devices so the function also requires the previous output.  For example, the C-element in figure \ref{fig:element0} with two input wires $w_1$ and $w_2$ is modeled by the function\\
\begin{figure}[h]
\centering
\includegraphics[width=0.3\textwidth]{element0}
\captionof{figure}{C-element}
\end{figure}
\[
f_{C-E}(w_1,w_2) =\left\{
  \begin{array}{@{}ll@{}}
    0, & \text{if}\ w_1=0 \wedge w_2=0\\
    1, & \text{if}\ w_1=1 \wedge w_2=1\\
    v, & \text{otherwise}

@@

Each element is in one of two states, excited or stable.  
\begin{definition}
An element with inputs $w_1$, $w_2$, output $v$ and function $f$ is {\em excited} if $v\neq f(w_1,w_2)$.  
\end{definition}
This definition can be extended to include elements with any number of inputs by replacing $w_1$ and $w_2$ with the inputs of a specific element.  A natural extension to an excited element is to consider an element that is not excited.
\begin{definition} 
An element is stable if the element is not excited.
\end{definition}
Then, a stable element must satisfy $v= f(w_1,w_2)$.  An example of an excited and stable AND gate is shown in figure \ref{fig:gates1}.  An example of an excited and stable C-element is shown in figure \ref{fig:element1}.  For C-elements, there are two situations in which the element is stable.  One is where the output is being actively driven to its function value, i.e. when both inputs are 0 or both inputs are 1.  The second is where the inputs are mismatched and the C-element is holding the previous output.  In the second case, the C-element is stable regardless of the output's value.

\begin{figure}[h]
\centering
\includegraphics[width=0.8\textwidth]{gates1}
\captionof{figure}[AND excited and stable]{a) the AND element is excited  b) an excited element fires to become stable}
\end{figure}
\begin{figure}[h]
\centering
\includegraphics[width=0.9\textwidth]{element1}
\captionof{figure}[C-element excited and stable]{a) the C-element is excited  b) two types of stable C-elements}
\end{figure}

@@

\chapter{Fault Tolerant Circuit Under Normal Operation and Single Transient Fault}
\chaptermark{Fault Tolerant Circuit}
\section{Traces and circuit equivalence}
We solve the problem of single transient faults by taking a semi-modular circuit (the original circuit) and applying a transformation to the circuit by duplicating all of its circuit elements.  Due to the asynchronous nature of how signals are processed, corresponding wires in the resulting half circuits may have different values.  To solve this problem we add C-elements to synchronize the two versions of the same signal.   An example of this transformation applied to the simple inverter circuit is shown in figure \ref{fig:dupschemeex}.  Both the duplication and the additional C-elements configuration are shown.  In the rest of this chapter we define a way to compare two circuits and to show their equivalence.  Then we propose our transformation method and show that under normal operation (no faults), the fault-tolerant circuit is equivalent to the original.  Finally we show that if a single transient fault occurs, one side of the fault-tolerant circuit is equivalent to the original and we can still extract the signals that we desire.\\
\begin{figure}
  \centering
    \includegraphics[width=.9\textwidth]{transf3Csimp}
  \caption[Fault-tolerant transformation of an example circuit]{Fault-tolerant transformation of an example circuit: a) original circuit b) after the fault-tolerant transformation}
  \label{fig:dupschemeex}
\end{figure}

@@

Another wrong approach for the fault-tolerant transformation is that if we take the following view for transitions, a transition of wire $w$ in the original circuit is complete if both $w^A$ and $w^B$ transitions in the fault-tolerant circuit, then some signals may transition out of order.  We use the earlier example of figure \ref{fig:counterexhalf} to show this as well.  Following the same transitions outlined above, $b^A$ and $b^B$ can finish their transitions before $a^B$ and $c^A$ transitions.  Then the $b+$ transition is actually completed before $a+$ or $c+$ and the combined trace no longer maps back to a valid trace in the original state graph.  Another approach we may take is to record the first transition of a wire in either half of the fault-tolerant circuit.  For example, if $w^A=w^B$ and then $w^A$ transitions, we map this to a transition on $w$ in the original circuit.  This approach ensures the resulting trace is a trace in the original circuit (can be shown by Lemma 1 and Lemma 2.1 in the next section), however, in the presence of faults, it is easy to see that it does not provide any fault tolerance.\\

>>>>

\chapter{Related Works}
\section{Fault Model}
A fault is defined as the physical difference between the good or correct system, and the current system.  In this work, the system is the asynchronous circuit.  An error, which is when the state of the system differs from the state of the good system and thus unable to perform some specified function, may be caused by a fault.  Hidden faults, where the fault does not cause an error, are also possible \cite{jha_gupta_2003}. \\

Faults can be split into permanent and non-permanent faults. Permanent faults are faults that affect the functional behavior of a system or circuit permanently.  Non-permanent faults are faults that affect the system's functional behavior for a finite, but unknown period of time.  Non-permanent faults can occur at random moments and are categorized into two types:  transient faults and intermittent faults \cite{jha_gupta_2003}.\\ %copied

Transient faults are caused by many environmental sources such as cosmic rays, $\alpha$-particles, pollution, humidity, temperature, pressure, vibration, power supply fluctuations, electromagnetic interference, static electrical discharges, and ground loops. In particular, $\alpha$-particles are a major cause of transient faults.  The $\alpha$-particles can interact with the semiconductor material which result in the generation of electron-hole pairs.  When an $\alpha$-particle collides and interacts with the circuit, it can induce a voltage change on a wire.  When the $\alpha$-particle causes a bit flip on a wire, this is called a single-event upset (SEU).  When the SEU is on the output wire of a gate, that wire will eventually return to its previous value.  If the SEU is on the output of a memory element, such as a flip flop, latch, or C-element, that faulty value is instead stored.  SEUs include both single-bit and multi-bit upsets, but single-bit upsets, where a single wire is affected, are most common.  There are other transient Single Event Effects (SEEs) such as the single-event transient (SET), where there is a voltage pulse on the wire.  SEUs are the most abundant type of transient faults and we focus on SEUs in this work.  \cite{jha_gupta_2003}\\

Intermittent faults, another non-permanent fault, are caused by non-environmental conditions such as loose connections, deteriorating or aging components, critical timing (hazards and race conditions), resistance and capacitance variations (which may lead to timing faults), physical irregularities, and noise.  Intermittent faults can transition into permanent faults naturally but may take a long time ranging from hours up to months. \\

Permanent faults are caused by radiation, incorrect connections between integrated circuits, broken components or parts of components, incorrect silicon and metal connections (manufacturing problem), and functional design errors. Most permanent faults occur during fabrication. Much work has been done on permanent faults and designing different tests at production time to uncover the presence and location of these faults. \cite{jha_gupta_2003} \cite{giz_book_2006}.  Unfortunately, these methods do not work for non-permanent faults as these faults affect the system for a shorter time period and thus may not be detected at all.  In addition to possible manufacturing problems, permanent faults can develop during operation due to radiation and aging effects.  For example, high energy ions can impact the transistors and the circuit, causing short-circuiting, rupture of the gate oxide insulation and other permanent faults.\\ %copied from SEE eetimes article

As mentioned before, we focus on SEUs as our main fault model because they are the most abundant transient faults.  Transient faults happen sporadically so there are usually long intervals of time, such as days to months, between consecutive SEUs.  Thus we model each fault as happening one at a time and call it a single transient fault.  We will not be concerned with intermittent faults because they take a while to develop.  We also ignore permanent faults because they are usually detected at production time.  \\

There are two main sources of radiation that cause SEUs: one is cosmic rays from space which interact with the atmosphere to produce high energy neutrons; another is the $\alpha$-particles caused by trace impurities of radioactive elements present in the packaging material of the chip and the silicon substrate. Radiation from cosmic rays become more severe at higher elevations. Thus, initially a lot of the work in fault-tolerant design was done for aviation and space flight.  However, fault-tolerant design is becoming more important as radiation effects become more prominent in new circuit technology.

@@

\section{Asynchronous Circuit and Synthesis}
In the previous chapter, we described how an asynchronous circuit works from a higher-level perspective; that it should work under varying delays. In this section, we will cover some definitions and properties of such a circuit and how the circuit is synthesized from a higher-level specification.  
%Both Speed Independent and Delay Insensitive technologies will be introduced. There are differences but also many similarities between the two methods. Even though this work is based on the Speed Independent model, there has been some previous work done on the Quasi-delay Insensitive model so I wanted to provide some context to make a direct comparison.

\subsection{Speed Independent Circuits}
Circuits are composed of logic elements and wires. The unbounded gate delay model assumes an arbitrary delay on the logic elements and no delay on wires.  This delay model is the basis for Muller circuits \cite{Muller_59}.  Circuits that operate correctly regardless of the delay are called Speed Independent Circuits. The designer specifies the correct operation of the circuit.  For example, given certain input values, the circuit will return specified output values. 
There are many ways to represent specifications such as with Petri nets and state graphs, and there are established methods to build asynchronous circuits that satisfy the specifications \cite{myers_book_2004}. We won't go into these representations and methods in detail, other than the aspects that are relevant to this work. We work with digital circuits, so the wires take on Boolean values of either `1' or `0'.\\

%Wires in the circuits take on Boolean values of either `1' or `0'. A state is defined by assigning a Boolean value to each wire of the circuit. States are unique in that no two states share the same wire values. There will be a formal definition of wires and states in the next chapter.

@@

%C-elements
\subsection{C-elements}
The C-element proposed by Muller, is an important building block for asynchronous circuits. It acts as a memory element because it can sometimes store a previous output value.  A C-element has two inputs, when both inputs are low, the C-element output goes low, and when both inputs are high, the C-element output goes high.  However, when the inputs are different Boolean values, the output gets held at its previous value. This can be achieved in a number of ways such as a gate-level implementation or a transistor-level implementation with feedback using a weak transistor, still other implementations use capacitance to achieve this weak feedback.  C-elements are an essential component of Muller circuits and asynchronous circuit designs. \\
\begin{figure}
\includegraphics[width=\textwidth]{celement}
\captionof{figure}[C-element: logic and implementation]{a. logic of the C-element. A, B are inputs, C is the output \\ b. diagram of a C-element at the gate level c. transistor-level implementation}
\end{figure}
%replace with figure from ...

For a semi-modular state graph, there are many existing methods of implementing a Speed-Independent asynchronous circuits that satisfies the state graph.  Generalized C-element implementation and standard C-implementation are two possible circuit implementations using C-elements.  
A generalized C-element directly implements switching logic and the C-element as one circuit element at the transistor-level \cite{myers_book_2004}.  This means, for the purpose of the asynchronous circuit, there is only one delay for this logic block.  The logic for the C-element component is as follows: there is a set condition, the condition for the output wire to transition to a high value; and a reset condition, the condition for the output wire to transition to a low value.  The set and reset conditions are designed to be mutually exclusive so that they are never both true at the same time.  The set and reset conditions are implemented with transistors and inserted into the C-element's pull-up and pull-down networks respectively.\\

%We skim the details in finding the set and reset conditions from the circuit specifications. For each condition set and reset, there is an ON-set, OFF-set and Don't Care set. ON-set are all input values where the wire is excited and a transition will occur. The OFF-set includes all input values where the wire cannot be excited to some value per the circuit specifications. Input values that are neither in the ON-set or OFF-set are in the Don't Care set. So, for the set condition, the ON-set are all input values where the wire must transition to high. The OFF-Set includes all input values where the wire is excited instead to low, as well as input values for which the output wire must remain low. The input values of the Don't Care set assign the output wire to be either high or low. It does not matter which value is assigned since it does not disturb the function of the circuit and only helps for circuit optimization. \\
We assume that the circuits we want to make fault-tolerant are synthesized according to these two implementations.

%The same procedure applies as well in finding the ON-set, OFF-set and Don't Care set for reset except substituting in reset for set and flipping the value of the wire from high to low and vice versa. The set and reset conditions constructed this way then follows the desired circuit specification. The set and reset conditions are implemented using gates and are combined into the output wire through a C-element. This entire block of gates and C-element for the output wire is called the gate logic. \\ %Beerel and Myers.... also expand?

%Even though the generalized C-element is presented as a C-element with set and reset as inputs, the transistor level implementation differs slightly. We assume a transistor level implementation of the set and reset combined with the C-element so that there is only one delay for the entire logic block. In addition, this is not a normal C-element because set is used exclusively for the pull-up logic and reset is used for the pull-down logic. In a normal C-element we allow for any four combinations of the values of the set and reset input wires. However, using the pull-up and pull-down logic setup, we do not allow for when both set and reset is `ON' as the output wire would simultaneously be connected to the high voltage and ground. Such an output would have indeterminable voltage values and violate the digital assumptions. Thus, the set and reset logic is designed so that they cannot overlap. This corresponds to $\overline{set}\vee\overline{reset} $ is always TRUE.\\
\begin{center}
\includegraphics[width=.7\textwidth]{genC}
\end{center}
\captionof{figure}[Generalized C-element implementation]{Transistor level implementation of a generalized C-element. The + signs are the set logic inputs and - signs are the reset logic inputs. For this particular element, set is in3 AND in4 while reset is in1 AND in2.}

%The main difference is that the set logic and the reset logic must be designed so that they do not turn `ON' at the same time.\\

The standard C-implementation is a gate-level implementation.  Each wire from the specification is implemented as a C-element with non-looping cascades of gates as the C-element's two inputs.  That is, each of the two inputs to the C-element is itself an output of a cascade of gates.  The C-element operates like a typical C-element where if the two inputs are `1', then the output is `1'; and when the two inputs are `0', the output is `0'; and when the two inputs have different values, the output value is held.  Additionally, Myers shows a method where the gate logic for the inputs to the C-element can always be re-implemented with two-input gates \cite{myers_book_2004}.  Since the C-element itself is also a two-input circuit element, then the entire circuit is composed of two-input circuit elements.  In such an implementation, each gate or C-element is assumed to have its own delay that is arbitrary in length.  Thus, this gate-level implementation allows for more generic design.
%explain sop?

\subsection{Quasi Delay Insensitive Circuits}
%QDI
Quasi Delay Insensitive Circuits (QDI) are related to Delay Insensitive Circuits. Previously we mentioned that Delay Insensitive circuits operate under the assumption that both the gate and the wire have unbounded gate delays. In particular each branch of a fork will have different delays and can be treated as separate wires. One result about Delay Insensitive circuits is that it is a very restricted class. Instead if that condition is relaxed so that some of the forks are allowed to be isochronic, then a more general class of circuits can be constructed. Note that in the case where all forks are isochronic, then this becomes the same as speed independent circuits. \\
The gates of a QDI circuit is formed from Production Rules (PRs). For the single output case, a gate takes a set of variables as inputs and outputs a variable. Again, a transition on the variable $x$ is denoted as $x\uparrow$, a transition from $x=0$ to $x=1$; and $x\downarrow$, a transition from $x=1$ to $x=0$. Then the two PRs associated with the gate with output $x$ is :
\begin{align*}
B_u\mapsto x\uparrow \\
B_d\mapsto x\downarrow 
\end{align*}
Where $B_u$ and $B_d$ are called guards. Each guard is a boolean condition formed by some combination of the input variables. When a guard is evaluated to be $True$ then the transition is activated. When it is $False$ then the corresponding transition does not happen. Note that similar to the generalized C-elements, both guards cannot be active at the same time. In addition, it is assumed that each guard is in a sum of products of the input variables. The PRs of a circuit are evaluated all at the same time. At any point in time, if a transition is activated, it will transition in a finite amount of time as long as the corresponding guard is held at $True$. This is due to the assumption of the arbitrary delay of a gate. Note that again, like the generalized C-element implementation, we assume that the transistor level implementation of a PR allows any arbitrary sum of products to be made where there are no delays within the logic of the sum of products.\\

@@

%soft error protection <- error correcting codes for SRAM memory
Because faults cannot be eliminated in digital circuits, different methodologies have been used to make the circuits fault-tolerant. Much of the research done in this area are for synchronous circuits. There has also been some research for asynchronous circuits using some similar techniques. First, there is an overview of some common techniques for synchronous circuits, then there is a summary of the current techniques for asynchronous circuits. Some of the fault-tolerant techniques in asynchronous design borrow or are inspired by the much more developed synchronous circuits field. However, due to the difference in timing assumptions of the two types of circuits the techniques are not completely transferable. We describe why some of the solutions from synchronous circuits cannot be directly applied to asynchronous circuits.\\

%- used in the early days for space and flight, and systems that CANNOT go offline/errors like stocks and banking.
%- different techniques... explain a bit about each
%-hardware, software, hybrid, 
%-detection vs correction
%- ECC/ parity checking (needs software/ controller) (Mitra's method)
%-TMR is still super important (possibly why TMR is not viable for async)
In the early days, there were certain systems for which eliminating faults and errors were very important. These include space and aviation control systems where due to the altitude there are more radiation and thus more faults. In addition, faults that cause error is very detrimental to these systems and may cause failures leading to loss of life. Other examples of systems requiring reliable circuits include banking and stock market applications where the system must function without error for hours at a time and in case of any error may have devastating consequences, such as a bit flip of the most significant bit in a money transaction. \\

@@

\chapter{Asynchronous Circuit Concepts \& Definitions}
\section{Assumptions:  Boolean signals}
We make the assumption that all signals are Boolean, that is, the signals are either `logic 0' or logic 1'.  When the gates and C-elements are excited, and their outputs transition, we assume that transition is instantaneous and that after the transition, the signals are still Boolean.   
Even in cases where the wires are affected by faults, we assume that the new value of the wires are Boolean, and if they are inputs into circuit elements, that the output of the circuit elements also remain Boolean.  For each circuit, there is a finite set of wires, which we call $W$.  An exact definition $W$ will be presented at a later section, but for now just think of $W$ as including all the wires of some circuit.

@@

For each element, there exists a function that maps the inputs into a binary value.  This function is represented by $f:[I \to \{0,1\}] \to \{0,1\}$ where $I$ is the set of input wires of the element, a subset of $W$.  If we assume a two input element, we name the input wires $w_1$ and $w_2$.  The function models the designed operation of the element such that if the element had no delay, $f(w_1,w_2)$ is the output value of the element.  With asynchronous delay, if the inputs remain constant, then the output will eventually take the value of $f(w_1,w_2)$.  For gates, such a function is easily defined.  For example, an AND gate such as in Figure \ref{fig:gates0} with two inputs $w_1$ and $w_2$ is modeled by the function 
\begin{figure}[h]
\centering
\includegraphics[width=0.4\textwidth]{gates0}
\captionof{figure}{AND gate}
\label{fig:gates0}
\end{figure}
\[
f_{AND}(w_1,w_2)= w_1\wedge w_2
\]

For C-elements, these are memory holding devices so the function also requires the previous output.  For example, the C-element in Figure \ref{fig:element0} with two input wires $w_1$ and $w_2$ is modeled by the function\\
\begin{figure}[h]
\centering
\includegraphics[width=0.3\textwidth]{element0}
\captionof{figure}{C-element}
\label{fig:element0}
\end{figure}
\[
f_{C-E}(w_1,w_2) =\left\{
  \begin{array}{@{}ll@{}}
    0, & \text{if}\ w_1=0 \wedge w_2=0\\
    1, & \text{if}\ w_1=1 \wedge w_2=1\\
    v, & \text{otherwise}

@@

Each element is in one of two states, excited or stable.  
\begin{definition}
An element with inputs $w_1$, $w_2$, output $v$ and function $f$ is {\em excited} if $v\neq f(w_1,w_2)$.  
\end{definition}
This definition can be extended to include elements with any number of inputs by replacing $w_1$ and $w_2$ with the inputs of a specific element.  A natural extension to an excited element is to consider an element that is not excited.
\begin{definition} 
An element is stable if the element is not excited.
\end{definition}
Then, a stable element must satisfy $v= f(w_1,w_2)$.  An example of an excited and stable AND gate is shown in Figure \ref{fig:gates1}.  An example of an excited and stable C-element is shown in Figure \ref{fig:element1}.  For C-elements, there are two situations in which the element is stable.  One is where the output is being actively driven to its function value, i.e. when both inputs are 0 or both inputs are 1.  The second is where the inputs are mismatched and the C-element is holding the previous output.  In the second case, the C-element is stable regardless of the output's value.

\begin{figure}[h]
\centering
\includegraphics[width=0.8\textwidth]{gates1}
\captionof{figure}[AND excited and stable]{a) the AND element is excited  b) an excited element fires to become stable}
\label{fig:gates1}
\end{figure}
\begin{figure}[h]
\centering
\includegraphics[width=0.9\textwidth]{element1}
\captionof{figure}[C-element excited and stable]{a) the C-element is excited  b) two types of stable C-elements}
\label{fig:element1}
\end{figure}

@@

\chapter{Fault Tolerant Circuit Under Normal Operation and Single Transient Fault}
\chaptermark{Fault Tolerant Circuit}
\section{Traces and circuit equivalence}
We solve the problem of single transient faults by taking a semi-modular circuit (the original circuit) and applying a transformation to the circuit by duplicating all of its circuit elements.  Due to the asynchronous nature of how signals are processed, corresponding wires in the resulting half circuits may have different values.  To solve this problem we add C-elements to synchronize the two versions of the same signal.   An example of this transformation applied to the simple inverter circuit is shown in Figure \ref{fig:dupschemeex}.  Both the duplication and the additional C-elements configuration are shown.  In the rest of this chapter we define a way to compare two circuits and to show their equivalence.  Then we propose our transformation method and show that under normal operation (no faults), the fault-tolerant circuit is equivalent to the original.  Finally we show that if a single transient fault occurs, one side of the fault-tolerant circuit is equivalent to the original and we can still extract the signals that we desire.\\
\begin{figure}
  \centering
    \includegraphics[width=.9\textwidth]{transf3Csimp}
  \caption[Fault-tolerant transformation of an example circuit]{Fault-tolerant transformation of an example circuit: a) original circuit b) after the fault-tolerant transformation}
  \label{fig:dupschemeex}
\end{figure}

@@

Another wrong approach for the fault-tolerant transformation is that if we take the following view for transitions, a transition of wire $w$ in the original circuit is complete if both $w^A$ and $w^B$ transitions in the fault-tolerant circuit, then some signals may transition out of order.  We use the earlier example of Figure \ref{fig:counterexhalf} to show this as well.  Following the same transitions outlined above, $b^A$ and $b^B$ can finish their transitions before $a^B$ and $c^A$ transitions.  Then the $b+$ transition is actually completed before $a+$ or $c+$ and the combined trace no longer maps back to a valid trace in the original state graph.  Another approach we may take is to record the first transition of a wire in either half of the fault-tolerant circuit.  For example, if $w^A=w^B$ and then $w^A$ transitions, we map this to a transition on $w$ in the original circuit.  This approach ensures the resulting trace is a trace in the original circuit (can be shown by Lemma 1 and Lemma 2.1 in the next section), however, in the presence of faults, it is easy to see that it does not provide any fault tolerance.\\
